---
title: "HW2 - Do Households Value Transit-rich Neighborhoods compared to Others in Boston?"
author: "Jun, Youngsang"
date: "September 19, 2024"
output: 
  html_document: 
    theme: readable
    code_folding: hide
editor_options: 
  markdown: 
---

*Since the Tremont Street subway opened in 1897 as North America’s first subway tunnel, Boston's subway system has become one of the most beloved public transportation options for its residents. As the Metropolitan Area Planning Council (MAPC) discovered that a disproportionate share of housing and jobs are located near transit in their study, particularly with the sharp increase in rents over the past decade, it is necessary to assess whether households value transit-rich neighborhoods compared to others in Boston and to prepare appropriate policy alternatives. To answer this question, this report conducted a Transit Oriented Development (TOD) analysis of Boston, by using subway stations data from the Massachusetts Bay Transportation Authority (MBTA), along with American Community Survey (ACS) data of 2012 and 2022. The result shows that people are willing to pay higher prices to live near a transit station in 2022 compared to 2012.*

## 1. Data Wrangling Work

The spatial analysis began with data wrangling by using Boston's subway arc and node data, which includes 5 lines and 153 stations, provided by MBTA, along with ACS data from 2012 and 2022 for 438 tracts within Suffolk County, provided by the U.S. Census Bureau. The result of data wrangling work is as follows:

a.  Load Libraries and Initial Setup

```{r setup, warning = FALSE, message = FALSE}
# Install and Load Libraries
library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot)
library(jtools)     # for regression model plots
library(broom)
library(tufte)
library(rmarkdown)
library(pander)
library(tidycensus)

# functions and data directory
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"

# Source for Multiple Ring Buffer
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

# Set Color Table
palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")

# Disable Scientific Notation
options(scipen=999)
options(tigris_class = "sf")

# Set ACS API Key
census_api_key("b2835f54d89a4499ba29829c908967b86765b345", overwrite = TRUE)
```

b.  Wrangle ACS Data of Suffolk County (which Boston seats of) 2012 and 2022

```{r getACS, warning = FALSE, message = FALSE, results = "hide"}
# Get ACS Data of Suffolk County 2012 and 2022
tracts12 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2012, state="MA",
          county="Suffolk", geometry=TRUE) %>% 
  st_transform('EPSG:26986') # Use NAD 83 CRS
tracts12 <- subset(tracts12, GEOID!='25025990101') # Error
tracts22 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), # Since B25026_001 is Total Population in 
                                        # Occupied Housing Units, replaced it with B01003_001E
          year=2020, state="PA",
          county="Philadelphia", geometry=TRUE) %>% 
  st_transform('EPSG:26986') # Use NAD 83 CRS

# Transforming Long Data to Wide Data Using Spread
tracts12 <- 
  tracts12 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

tracts22 <- 
  tracts22 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

# Create Column that Calculates Proportion of Whites, Bachelors and Poverty
tracts12 <- 
  tracts12 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2012") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

tracts22 <- 
  tracts22 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2022") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

# Row Bind of 2012 and 2022
allTracts <- rbind(tracts12,tracts22)
```

c.  Wrangling Boston's Subway Arc and Node Data (MBTA Data)

```{r get_MBTA_data, warning = FALSE, message = FALSE, results = "hide"}
# Get MBTA Data of Transit Line and Node
MBTA_ARC <- st_read('./MBTA_RT/MBTA_ARC.shp')
MBTA_NODE <- st_read('./MBTA_RT/MBTA_NODE.shp')

# Create a Column in MBTA_NODE to Distinguish Transfer Stations
MBTA_NODE <- MBTA_NODE %>%
  mutate(transferstation = ifelse(LINE == "BLUE", "BLUE",ifelse(LINE == "ORANGE", "ORANGE",ifelse(LINE == "RED", "RED",ifelse(LINE == "SILVER", "SILVER", ifelse(LINE == "GREEN", "GREEN","TRANSFER")))))) %>%
  mutate(IDD=1:n())
```

d.  Check if the Locations of MBTA Data are precise in Online map (via Mapview)

```{r mapViewMBTA, warning = FALSE, message = FALSE}
mapview(tracts22)+mapview(MBTA_ARC)+mapview(MBTA_NODE)
```

e.  Buffer Relating MBTA Data and Set TOD/Non-TOD area

In this study, TOD area is defined as the total area of tracts whose centroids are located in 0.5 miles of each transit station. The following figure shows the buffer within 0.5 mile of each transit station. The color of line represents each service line, and black dots denote transfer stations.

```{r buffer, cache = TRUE, warning = FALSE, message = FALSE}
stopBuffer <- st_buffer(MBTA_NODE, 805) # 805 meters which is equivalent to 0.5 miles
stopUnion <- st_union(st_buffer(MBTA_NODE, 805))
stopUnion <- st_make_valid(stopUnion) 
MBTABuffers <- 
  rbind(
     stopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     stopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, 
          aes(color = LINE), 
          show.legend = "line", size= 2) +
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E")) +
  geom_sf(data=MBTA_NODE,
          aes(color = transferstation), 
          show.legend = "point", size= 1)+
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E", "black"), labels = c("BLUE", "GREEN", "ORANGE", "RED", "SILVER", "TRANSFER STN")) +
  labs(title="Buffer within 0.5 mile of Each Transit Stn", 
       subtitle="The color of line represents each line, and black dots denote transfer stations", 
       caption="Data: U.S. Census Bureau, MBTA") +
  mapTheme()
```

f.  Create Centroid

```{r centroids, warning = FALSE, message = FALSE}
buffer <- filter(MBTABuffers, Legend=="Unioned Buffer")

selectCentroids22 <-
  st_centroid(tracts22)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts22, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2022")
selectCentroids22_union <- st_union(selectCentroids22)%>%
  st_sf() %>%
  mutate(year = "2022")
selectCentroids12 <-
  st_centroid(tracts12)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts12, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2012")
selectCentroids12_union <- st_union(selectCentroids12)%>%
  st_sf() %>%
  mutate(year = "2012")
intersections1222_union <- rbind(selectCentroids22_union, selectCentroids12_union)

NODE22 <- st_intersection(MBTA_NODE, tracts22) %>%
  mutate(year = "2022")
NODE12 <- st_intersection(MBTA_NODE, tracts12) %>%
  mutate(year = "2012")
MBTA_NODE_CLIP <-
  rbind(NODE22,NODE12)
```

g.  Row Bind for TOD, Non-TOD, by year

```{r rbind_TOD, warning = FALSE, message = FALSE}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2012", MedRent * 1.2747, MedRent)) %>% # Cumulative price change Inflation from 2012 to 2022 is	27.47%
  mutate(tractArea=st_area(geometry)) %>%
  mutate(tractpopDens=TotalPop*2589988/units::drop_units(tractArea))
```

## 2. Four small-multiple visualizations comparing four selected Census variables across time and space (TOD vs. non-TOD)

a.  Population Density by Tract, 2012-2022

The figure below shows the population density per tract and TOD area in Boston. The region within the red line, representing the TOD area, is colored green, while the Non-TOD area, farther from the station, is shown in dark blue. This shows that population density tends to be higher near transit stations. In particular, the tract with the highest population density in 2012 was located between Symphony, Hynes Convention, and Northeastern stations, with a density of 117,000 residents per square mile. By 2022, the population density in this area decreased to 87,000 residents per square mile.

```{r TODmap1, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=tractpopDens))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Population Density by Tract, 2012-2022",
    subtitle = "Person/mi^2; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

b.  Median Monthly Rent by Tract, 2012-2022

The following figure demonstrates the Median Monthly Rent per tract in Boston. The colors near the station are light green or yellow, and the colors far from the station are dark green, showing that the Median Rent tends to be higher in the TOD area overall. In particular, when comparing 2012 and 2022, the Median Rent in the TOD area increased to $3,000 in some tracts in 2022. Note that the Median Rent in 2012 was displayed considering inflation rate from 2012 to 2022 (27.47%) in order to derive reasonable results.

```{r TODmap2, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=MedRent.inf))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Median Monthly Rent, 2012-2022",
    subtitle = "Real dollars; The red border denotes areas close to subway stns;\nThe black dots denote subway stns\nInflation rate considered (compared Rent in 2022 to Rent in 2012*1.2747)",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

c.  Proportion of White by Tract, 2012-2022

The following figure shows the proportion of white population by tract in Boston. In the Boston area and the northern and southwestern parts of Suffolk County, the difference in white population between TOD and non-TOD areas was not large, while in the central and southern parts of Suffolk County, the difference in white population between TOD and non-TOD areas was large.

```{r TODmap3, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctWhite))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of White, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

d.  Proportion of Bachelors by Tract, 2012-2022

The figure below shows that the proportion of Bachelors in TOD areas is higher than in non-TOD areas. The proportion of Bachelors in some tracts near subway stations in TOD areas is 10-15%, which is higher than the proportion of Bachelors in the rest of the areas, which is less than 5%. Notably, Chestnut Hill and South Boston areas have decreased in the percentage of college graduates in 2022 compared to 2012. There was no significant difference in Bachelors by year in other regions.

```{r TODmap4, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctBachelors))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of Bachelors, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

e.  Proportion of Poverty by Tract, 2012-2022

The figure below shows that the proportion of Poverty in TOD areas is higher than in non-TOD areas. The proportion of Poverty in some tracts near subway stations in TOD areas is higher than the rest of the areas. There was no significant difference in Poverty by year in other regions.

```{r TODmap5, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctPoverty))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of Poverty, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

## 3. Grouped Bar Plot

This grouped bar plots denote that it has become more difficult to find a house for the same amount of money despite the decrease in population per tract, and this trend appears to be greater in TOD areas than in non-TOD areas. The average values of the proportion of white population, bachelors, and poverty per tract in 2012 and 2022 divided into TOD and Non-TOD areas are as follows. Those values are for reference only as it is an average of proportions and not an average per population or area of tracts.

```{r summary_barplot, warning = FALSE, message = FALSE}
allTracts.Summary <- 
  st_drop_geometry(allTracts.group) %>%
  group_by(year, TOD) %>%
  summarize(Rent = mean(MedRent.inf, na.rm = T),
            Population = mean(TotalPop, na.rm = T),
            Percent_White = mean(pctWhite, na.rm = T),
            Percent_Bach = mean(pctBachelors, na.rm = T),
            Percent_Poverty = mean(pctPoverty, na.rm = T))

allTracts.Summary %>%
  gather(Variable, Value, -year, -TOD) %>%
  mutate(Variable = factor(Variable, levels = c("Population", "Rent", "Percent_White", "Percent_Bach", "Percent_Poverty"))) %>%
  ggplot(aes(year, Value, fill= TOD ))+
    geom_bar(stat="identity", position="dodge")+
    facet_wrap(~Variable, scales="free", ncol=5)+
    scale_fill_manual(values=c("#bae4bc","#0868ac"))+
    labs(title="Indicator Differences across Time and Space")+
    plotTheme() + theme(legend.position="bottom",
                        plot.title = element_text(size=15), 
      axis.title = element_text(size=11), 
      axis.text = element_text(size=11),  
      strip.text = element_text(size=8)  # facet 제목 글씨 크기
    )
```

## 4. TOD Indicator Table per Tract

In 2022, the average population per tract in both TOD and Non-TOD areas decreased by less than 10% compared to 2012, but the average Median Rent per tract increased by 23.8% (57.8% when excluding inflation) in Non-TOD areas and 30.7% (66.5% when excluding inflation) in TOD areas. The average values of the proportion of white population, bachelors, and poverty per tract in 2012 and 2022 divided into TOD and Non-TOD areas are as follows. Those values are for reference only as it is an average of proportions and not an average per population or area of tracts.

```{r summary_table, warning = FALSE, message = FALSE}
allTracts.Summary %>%
  unite(year.TOD, year, TOD, sep = ": ", remove=T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Variable = factor(Variable, levels = c("Population", "Rent", "Percent_White", "Percent_Bach", "Percent_Poverty")))%>%
  mutate(Value=round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kbl(caption = "TOD Indicator Table per Tract in 2012 and 2022") %>%
  kable_styling(position="center") %>%
  kable_classic(full_width = T, html_font = "Arial")
```

## 5. A Graduated Symbol Map of Population and Median Rent within 0.5 mile of Each Transit Stn

A graduated symbol map showing the population and median rent within a 0.5-mile buffer around each transit station provides useful information about the trends of population and median rent near transit stations. To do this, I firstly iterated to intersect over the tracts with transit station buffer circles and to calculate the area of pieces of tracts in each transit station buffer. (The results of 2022 and 2012 are saved in `result22` and `result12`, respectively.) Then to get the total population in each transit station buffer, I multiplied the population density (`popDens`) of each tract by the area of the cut tracts, and summed the partial population (`partVal_pop`) within each buffer circle. To get a weighted mean of the median rent of each transit station buffer, I multiplied the median rent by the area of the cut tracts and divided it by the area of the transit station buffer, which is 2,034,900 square meters. Finally, I summed the partial median rent (`partVal_med`) within each buffer circle.

```{r pop_medRent_in_buffer, warning = FALSE, message = FALSE, results = "hide"}
# Calculate Population and Rent Density per Tract
allTracts_Dens <- allTracts %>%
  mutate(tractArea=st_area(geometry)) %>%
  mutate(popDens=TotalPop/tractArea)

allTracts_Dens22 <- subset(allTracts_Dens, year!='2012')
allTracts_Dens12 <- subset(allTracts_Dens, year!='2022')

# Calculate Area of Each Buffer
stopBuffer <- stopBuffer %>%
  mutate(bufferArea=st_area(geometry)) 

# Intersect (To Cut tracts22_Dens by stopBuffer12, 
# Loop through each stop and intersect with tracts22_Dens/tracts12_Dens)
results22 <- data.frame()
results12 <- data.frame()
stop_intersect <- data.frame()
for(i in 1:nrow(stopBuffer)) {
  stop_intersect <- st_intersection(stopBuffer[i,], allTracts_Dens22) %>% 
    mutate(partArea=st_area(geometry)) %>% 
    mutate(partPop = as.numeric(partArea*popDens)) %>% 
    mutate(partMed = as.numeric(partArea*MedRent/2034900))
  results22 <- rbind(results22, stop_intersect)
  stop_intersect <- data.frame()
  print(i)
}

for(i in 1:nrow(stopBuffer)) {
  stop_intersect <- st_intersection(stopBuffer[i,], allTracts_Dens12) %>% 
    mutate(partArea=st_area(geometry)) %>% 
    mutate(partPop = as.numeric(partArea*popDens)) %>% 
    mutate(partMed = as.numeric(partArea*MedRent/2034900))
  results12 <- rbind(results12, stop_intersect)
  stop_intersect <- data.frame()
  print(i)
}
```

```{r pop_medRent_in_buffer2, warning = FALSE, message = FALSE}
a <- st_drop_geometry(results22) %>%
  group_by(IDD, year) %>%
  summarize(
    partVal_pop=sum(partPop, na.rm=TRUE),
    partVal_med=sum(partMed, na.rm=TRUE)
  )

b <- st_drop_geometry(results12) %>%
  group_by(IDD, year) %>%
  summarize(
    partVal_pop=sum(partPop, na.rm=TRUE),
    partVal_med=sum(partMed, na.rm=TRUE)
  )

abGraph <- rbind(a,b)
abGraph <- left_join(abGraph, MBTA_NODE, by="IDD")
#medGraph <- rbind(b,d) 
#medGraph <- left_join(medGraph, MBTA_NODE, by="IDD")
```

### 5.1. Map of Population within 0.5 mile of Each Transit Stn

This graduated symbol map shows that the total population within 0.5 miles of each transit station. The most populated transit station within a 0.5 mile radius is Massachusetts Avenue, with 33,000 residents, both in 2022 and 2012. There is no significant difference in total population within 0.5 miles of each transit station between 2022 and 2012.

```{r PopRentPlot_12-22, warning = FALSE, message = FALSE}
#PLOT for Population 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=abGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal_pop), size = as.numeric(partVal_pop), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="C", labels = comma)+
  scale_size_continuous(range=c(2,8), labels = comma)+
  labs(title="Population within 0.5 mile of Each Transit Stn", 
       subtitle="Boston (Suffolk County), MA", 
       caption="Data: U.S. Census Bureau, MBTA") +
  guides(color = guide_legend("Population (residents)"), size = guide_legend("Population (residents)")) +
  theme_void()+
  theme(legend.position = "right")
```

### 5.2. Map of Median Rent within 0.5 mile of Each Transit Stn

This graduated symbol map shows that the average median rent within 0.5 miles of each transit station. The highest median rent transit stations within 0.5-mile radius in 2022 and 2012 were South Station (average 2,788 dollars) and Haymarket Station (average 1,831 dollars), respectively.

```{r PopRentPlot_12-22_2, warning = FALSE, message = FALSE}
#PLOT for Median Rent 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=abGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal_med), size = as.numeric(partVal_med), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="D", labels = comma)+
  scale_size_continuous(range=c(2,10), labels = comma)+
  labs(title="Median Rent within 0.5 mile of Each Transit Stn", 
       subtitle="Boston (Suffolk County), MA", 
       caption="Data: U.S. Census Bureau, MBTA") +
  guides(color = guide_legend("Median Rent ($)"), size = guide_legend("Median Rent ($)")) +
  theme_void()+
  theme(legend.position = "right")
```

## 6. Mean Rent as a Function of Distance to Subway Stations

The color of each tract on the following map indicates the distance from the nearest subway station in 0.5-mile intervals for 2012 and 2022. Most TOD areas are outlined in red, and most TOD areas are depicted in dark blue which means that they are located within 0.5 miles of a subway station. Tracts farther away from subway stations are shown in yellow.

```{r medRent_Dist_Relation, warning = FALSE, message = FALSE}
allTracts.rings <-
  st_join(
    st_centroid(dplyr::select(allTracts, GEOID, year)),
    multipleRingBuffer(st_union(MBTA_NODE), 14484, 805)) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year),
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance=distance/1609)

ggplot() +
  geom_sf(data=allTracts.rings, aes(fill=distance))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", color='white', size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Distance to Subway Stns, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe white dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

The graph below illustrates that median rents are overall higher in areas closer to a subway station. In particular, up to 1.5 miles, median rents tend to increase as the distance to a station decreases. However, there was no significant difference in median rents between areas more than 2 miles from a subway station in both 2012 and 2022. Seeing median rents difference by year, between within 0.5 miles of a station and those 1 mile or farther away in 2022 were greater than in 2012. However, median rents difference in the 2.0-2.5 mile range were higher than those in the 2.5-mile and 0.5-2.0 mile ranges in both year.

```{r medRent_Dist_Relation2, warning = FALSE, message = FALSE}
MedRent_sum22 <- allTracts.rings %>%
  filter(year %in% "2022", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2022)

MedRent_sum12 <- allTracts.rings %>%
  filter(year %in% "2012", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2012)

MedRent_sum <- rbind(MedRent_sum22,MedRent_sum12)


#MedRent_mean <- aggregate(MedRent ~ distance, data = allTracts.rings, FUN = mean, na.rm=TRUE, na.action=NULL)
MedRent_sum$year <- as.factor(MedRent_sum$year)
ggplot(MedRent_sum, aes(x = distance, y = mean_value, group=year, color=year)) +
  geom_line() +
  scale_x_continuous(limits = c(0.5, 3.6), breaks = seq(0.5, 3.6, by = 0.5)) +
  scale_y_continuous(limits = c(0, 2000), breaks = seq(0, 2000, by = 200)) +
  scale_color_manual(values = c("2022" = "red", "2012" = "blue"))+
labs(
    title = "Median Monthly Rent as a Function of Distance from Subway Stns",
    subtitle = "Boston (Suffolk County), MA",
    caption = "Data: U.S. Census Bureau, MBTA",
    x="Distance from Subway Stns (Miles)", 
    y="Median Monthly Rent ($)",
    color = "Year") +
  theme_minimal()
```

# Conclusion:

In 2022, it has become more difficult to find a house for the same amount of money in Boston compared to 2012, despite the decrease in population per tract. This trend appears to be greater in TOD areas than in non-TOD areas. Nevertheless, as the graduated symbol maps indicate, there is no significant difference in total population within 0.5 miles of each transit station between 2022 and 2012. This means that in 2022, people are willing to pay higher prices to live near a transit station compared to 2012. However, it is important to note that this analysis does not establish a causal relationship between these findings and TOD development. Therefore, I recommend that the City Council representatives of Boston make city of Boston and MBTA continue their current TOD initiatives, while also monitoring whether these developments align with the goals of addressing Boston’s housing crisis and improving access to public transportation.
