---
title: "HW2 - Do Households Value Transit-rich Neighborhoods compared to Others in Boston?"
author: "Jun, Youngsang"
date: "September 19, 2024"
output: 
  html_document: 
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---

*Since the Tremont Street subway opened in 1897 as North America’s first
subway tunnel, Boston's subway system has become one of the most beloved
public transportation options for its residents. As the Metropolitan
Area Planning Council (MAPC) discovered that a disproportionate share of
housing and jobs are located near transit in their study, particularly
with the sharp increase in rents over the past decade, it is necessary
to assess whether households value transit-rich neighborhoods compared
to others in Boston and to prepare appropriate policy alternatives. To
answer this question, this report conducted a Transit Oriented
Development (TOD) analysis of Boston, by using subway sations data from
the Massachusetts Bay Transportation Authority (MBTA), along with
American Community Survey (ACS) data of 2012 and 2022. The result shows
\~\~\~\~\~\~\~ conclusions about willingness to pay for TOD.*

## Key Facts

1.  In 2022, it has become more difficult to find a house for the same
    amount of money in Boston compared to 2012, despite the decrease in
    population per tract. This trend appears to be greater in TOD areas
    than in non-TOD areas.
2.  
3.  

## 1. Data Wrangling Work

The spatial analysis began with data wrangling by using Boston's subway
arc and node data, which includes 5 lines and 153 stations, provided by
MBTA, along with ACS data from 2012 and 2022 for 438 tracts within
Suffolk County, provided by the U.S. Census Bureau. The result of data
wrangling work is as follows:

a.  Load Libraries and Initial Setup

```{r setup, warning = FALSE, message = FALSE}
# Install and Load Libraries
# install.packages('mapview')
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(mapview)
library(dplyr)
library(viridis)

# Disable Scientific Notation
options(scipen=999)
options(tigris_class = "sf")

# Source for Multiple Ring Buffer
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

# Set Color Table
palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")

# Set ACS API Key
census_api_key("b2835f54d89a4499ba29829c908967b86765b345", overwrite = TRUE)
```

b.  Wrangle ACS Data of Suffolk County (which Boston seats of) 2012 and
    2022

```{r getACS, warning = FALSE, message = FALSE, results = "hide"}
# Get ACS Data of Suffolk County 2012 and 2022
tracts12 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2012, state="MA",
          county="Suffolk", geometry=TRUE) %>% 
  st_transform('EPSG:26986') # Use NAD 83 CRS
tracts12 <- subset(tracts12, GEOID!='25025990101') # Error
tracts22 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), # Since B25026_001 is Total Population in 
                                        # Occupied Housing Units, replaced it with B01003_001E
          year=2022, state="MA",
          county="Suffolk", geometry=TRUE) %>% 
  st_transform('EPSG:26986') # Use NAD 83 CRS

# Transforming Long Data to Wide Data Using Spread
tracts12 <- 
  tracts12 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

tracts22 <- 
  tracts22 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

# Create Column that Calculates Proportion of Whites, Bachelors and Poverty
tracts12 <- 
  tracts12 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2012") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

tracts22 <- 
  tracts22 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2022") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

# Row Bind of 2012 and 2022
allTracts <- rbind(tracts12,tracts22)
```

c.  Wrangling Boston's Subway Arc and Node Data (MBTA Data)

```{r get_MBTA_data, warning = FALSE, message = FALSE, results = "hide"}
# Get MBTA Data of Transit Line and Node
MBTA_ARC <- st_read('./MBTA_RT/MBTA_ARC.shp')
MBTA_NODE <- st_read('./MBTA_RT/MBTA_NODE.shp')

# Create a Column in MBTA_NODE to Distinguish Transfer Stations
MBTA_NODE <- MBTA_NODE %>%
  mutate(transferstation = ifelse(LINE == "BLUE", "BLUE",ifelse(LINE == "ORANGE", "ORANGE",ifelse(LINE == "RED", "RED",ifelse(LINE == "SILVER", "SILVER", ifelse(LINE == "GREEN", "GREEN","TRANSFER"))))))
```

d.  Check if the Locations of MBTA Data are precise in Online map (via
    Mapview)

```{r mapViewMBTA, warning = FALSE, message = FALSE}
mapview(tracts22)+mapview(MBTA_ARC)+mapview(MBTA_NODE)
```

e.  Buffer Relating MBTA Data and Set TOD/Non-TOD area

In this study, TOD area is defined as the total area of tracts whose
centroids are located in 0.5 miles of each transit station. The
following figure shows the buffer within 0.5 mile of each transit
station. The color of line represents each service line, and black dots
denote transfer stations.

```{r buffer, cache = TRUE, warning = FALSE, message = FALSE}
stopBuffer <- st_buffer(MBTA_NODE, 805) # 805 meters which is equivalent to 0.5 miles
stopUnion <- st_union(st_buffer(MBTA_NODE, 805))
stopUnion <- st_make_valid(stopUnion) 
MBTABuffers <- 
  rbind(
     stopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     stopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, 
          aes(color = LINE), 
          show.legend = "line", size= 2) +
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E")) +
  geom_sf(data=MBTA_NODE,
          aes(color = transferstation), 
          show.legend = "point", size= 1)+
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E", "black"), labels = c("BLUE", "GREEN", "ORANGE", "RED", "SILVER", "TRANSFER STN")) +
  labs(title="Buffer within 0.5 mile of Each Transit Stn", 
       subtitle="The color of line represents each line, and black dots denote transfer stations", 
       caption="Data: U.S. Census Bureau, MBTA") +
  mapTheme()
```

f.  Create Centroid

```{r centroids, warning = FALSE, message = FALSE}
buffer <- filter(MBTABuffers, Legend=="Unioned Buffer")

selectCentroids22 <-
  st_centroid(tracts22)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts22, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2022")
selectCentroids22_union <- st_union(selectCentroids22)%>%
  st_sf() %>%
  mutate(year = "2022")
selectCentroids12 <-
  st_centroid(tracts12)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts12, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2012")
selectCentroids12_union <- st_union(selectCentroids12)%>%
  st_sf() %>%
  mutate(year = "2012")
intersections1222_union <- rbind(selectCentroids22_union, selectCentroids12_union)

NODE22 <- st_intersection(MBTA_NODE, tracts22) %>%
  mutate(year = "2022")
NODE12 <- st_intersection(MBTA_NODE, tracts12) %>%
  mutate(year = "2012")
MBTA_NODE_CLIP <-
  rbind(NODE22,NODE12)
```

g.  Row Bind for TOD, Non-TOD, by year

```{r rbind_TOD, warning = FALSE, message = FALSE}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2012", MedRent * 1.2747, MedRent)) %>% # Cumulative price change Inflation from 2012 to 2022 is	27.47%
  mutate(tractArea=st_area(geometry)) %>%
  mutate(tractpopDens=TotalPop*2589988/units::drop_units(tractArea))
```

## 2. Four small-multiple visualizations comparing four selected Census variables across time and space (TOD vs. non-TOD)

a.  Population Density by Tract, 2012-2022

The figure below shows the population density per tract and TOD area in
Boston. The region within the red line, representing the TOD area, is
colored green, while the Non-TOD area, farther from the station, is
shown in dark blue. This shows that population density tends to be
higher near transit stations. In particular, the tract with the highest
population density in 2012 was located between Symphony, Hynes
Convention, and Northeastern stations, with a density of 117,000 people
per square mile. By 2022, the population density in this area is
projected to decrease to 87,000 people per square mile.

```{r TODmap1, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=tractpopDens))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Population Density by Tract, 2012-2022",
    subtitle = "Person/mi^2; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

b.  Median Monthly Rent by Tract, 2012-2022

The following figure demonstrates the Median Monthly Rent per tract in
Boston. The colors near the station are light green or yellow, and the
colors far from the station are dark green, showing that the Median Rent
tends to be higher in the TOD area overall. In particular, when
comparing 2012 and 2022, the Median Rent in the TOD area increased by
more than \$3,000 in some tracts in 2022. Note that the Median Rent in
2012 was displayed considering inflation rate from 2012 to 2022 (27.47%)
in order to derive reasonable results.

```{r TODmap2, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=MedRent.inf))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Median Monthly Rent, 2012-2022",
    subtitle = "Real dollars; The red border denotes areas close to subway stns;\nThe black dots denote subway stns\nInflation rate considered (compared Rent in 2022 to Rent in 2012*1.2747)",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

c.  Proportion of White by Tract, 2012-2022

The following figure shows the proportion of white population by tract
in Boston. In the Boston area and the northern and southwestern parts of
Suffolk County, the difference in white population between TOD and
non-TOD areas was not large, while in the central and southern parts of
Suffolk County, the difference in white population between TOD and
non-TOD areas was large.

```{r TODmap3, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctWhite))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of White, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

d.  Proportion of Bachelors by Tract, 2012-2022

The figure below shows that the proportion of Bachelors in TOD areas is
higher than in non-TOD areas. The proportion of Bachelors in some tracts
near subway stations in TOD areas is 10-15%, which is higher than the
proportion of Bachelors in the rest of the areas, which is less than 5%.
Notably, Chestnut Hill and South Boston areas have decreased in the
percentage of college graduates in 2022 compared to 2012. There was no
significant difference in Bachelors by year in other regions.

```{r TODmap4, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctBachelors))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of Bachelors, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

e.  Proportion of Poverty by Tract, 2012-2022

The figure below shows that the proportion of Poverty in TOD areas is
higher than in non-TOD areas. The proportion of Poverty in some tracts
near subway stations in TOD areas is higher than the rest of the areas.
There was no significant difference in Poverty by year in other regions.

```{r TODmap5, warning=FALSE, message = FALSE}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctPoverty))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Proportion of Poverty, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

## 3. Grouped Bar Plot

This grouped bar plots denote that it has become more difficult to find
a house for the same amount of money despite the decrease in population
per tract, and this trend appears to be greater in TOD areas than in
non-TOD areas. The average values of the proportion of white population,
bachelors, and poverty per tract in 2012 and 2022 divided into TOD and
Non-TOD areas are as follows. Those values are for reference only as it
is an average of proportions and not an average per population or area
of tracts.

```{r summary_barplot, warning = FALSE, message = FALSE}
allTracts.Summary <- 
  st_drop_geometry(allTracts.group) %>%
  group_by(year, TOD) %>%
  summarize(Rent = mean(MedRent.inf, na.rm = T),
            Population = mean(TotalPop, na.rm = T),
            Percent_White = mean(pctWhite, na.rm = T),
            Percent_Bach = mean(pctBachelors, na.rm = T),
            Percent_Poverty = mean(pctPoverty, na.rm = T))

allTracts.Summary %>%
  gather(Variable, Value, -year, -TOD) %>%
  mutate(Variable = factor(Variable, levels = c("Population", "Rent", "Percent_White", "Percent_Bach", "Percent_Poverty"))) %>%
  ggplot(aes(year, Value, fill= TOD ))+
    geom_bar(stat="identity", position="dodge")+
    facet_wrap(~Variable, scales="free", ncol=5)+
    scale_fill_manual(values=c("#bae4bc","#0868ac"))+
    labs(title="Indicator Differences across Time and Space")+
    plotTheme() + theme(legend.position="bottom",
                        plot.title = element_text(size=15), 
      axis.title = element_text(size=11), 
      axis.text = element_text(size=11),  
      strip.text = element_text(size=8)  # facet 제목 글씨 크기
    )
```

## 4. TOD Indicator Table per Tract

In 2022, the average population per tract in both TOD and Non-TOD areas
decreased by less than 10% compared to 2012, but the average Median Rent
per tract increased by 23.8% (57.8% when excluding inflation) in Non-TOD
areas and 30.7% (66.5% when excluding inflation) in TOD areas. The
average values of the proportion of white population, bachelors, and
poverty per tract in 2012 and 2022 divided into TOD and Non-TOD areas
are as follows. Those values are for reference only as it is an average
of proportions and not an average per population or area of tracts.

```{r summary_table, warning = FALSE, message = FALSE}
allTracts.Summary %>%
  unite(year.TOD, year, TOD, sep = ": ", remove=T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Variable = factor(Variable, levels = c("Population", "Rent", "Percent_White", "Percent_Bach", "Percent_Poverty")))%>%
  mutate(Value=round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kbl(caption = "TOD Indicator Table per Tract in 2012 and 2022") %>%
  kable_styling(position="center") %>%
  kable_classic(full_width = T, html_font = "Arial")
```

## 5. A Graduated Symbol Map of Population and Median Rent within 0.5 mile of Each Transit Stn
A graduated symbol map showing data within a 0.5-mile buffer around each transit station provides useful information into the trends of population and median rent near transit stations. To do this, I calculated the population and median rent density per tract, which were derived from ACS data. Second, I iterated to intersect over the tracts with transit station buffer circles, and to multiple by population and median rent density and area of the cut tracts. Finally, I summed them within each buffer circle.

```{r pop_medRent_in_buffer, warning = FALSE, message = FALSE, results = "hide"}
# Calculate Population and Rent Density per Tract
allTracts_Dens <- allTracts %>%
  mutate(tractArea=st_area(geometry)) %>%
  mutate(popDens=TotalPop/tractArea) %>%
  mutate(medDens=MedRent/tractArea)

# Calculate Area of Each Buffer
stopBuffer <- stopBuffer %>%
  mutate(bufferArea=st_area(geometry)) 

# Intersect (To Cut tracts22_Dens by stopBuffer12, 
# Loop through each stop and intersect with tracts22_Dens/tracts12_Dens)
stop_results <- data.frame()
hap <- data.frame()
for(i in 1:nrow(stopBuffer)) {
  stop_intersect <- st_intersection(stopBuffer[i,], allTracts_Dens) %>% 
    mutate(partArea=st_area(geometry)) %>% 
    mutate(partPop = partArea*popDens) %>% 
    mutate(partMed = partArea*medDens)
  stop_results <- rbind(stop_results, stop_intersect)
  print(i)
}

hapPop22 <- stop_results %>%
  filter(year %in% "2022") 
hapMed22 <- stop_results %>%
  filter(year %in% "2022") 
hapPop12 <- stop_results %>%
  filter(year %in% "2012")
hapMed12 <- stop_results %>%
  filter(year %in% "2012") 

a <- aggregate(hapPop22$partPop, by=list(hapPop22$STATION), FUN=sum, na.rm=TRUE, na.action=NULL) 
colnames(a) <- c("STATION","partVal") 
a <- a %>%
  mutate(year="2022") %>%
  mutate(classification="Population by Subway Stn")
  
b <-aggregate(hapMed22$partMed, by=list(hapMed22$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(b) <- c("STATION","partVal")
b <- b %>%
  mutate(year="2022") %>%
  mutate(classification="Median Rent by Subway Stn")

c <-aggregate(hapPop12$partPop, by=list(hapPop12$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(c) <- c("STATION","partVal") 
c <- c %>%
  mutate(year="2012") %>%
  mutate(classification="Population by Subway Stn")

d <-  aggregate(hapMed12$partMed, by=list(hapMed12$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(d) <- c("STATION","partVal")
d <- d %>%
  mutate(year="2012") %>%
  mutate(classification="Median Rent by Subway Stn")

popGraph <- rbind(a,c)
popGraph <- left_join(popGraph,MBTA_NODE, by="STATION")
medGraph <- rbind(b,d) 
medGraph <- left_join(medGraph,MBTA_NODE, by="STATION")
```

### 5.1. Map of Population

This graduated symbol map shows that population within 0.5 mile of each transit station. Central Boston has upto 80,000 people.

```{r PopRentPlot_12-22, warning = FALSE, message = FALSE}
#PLOT for Population 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=popGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal), size = as.numeric(partVal), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="C")+
  scale_size_continuous(range=c(2,18))+
  labs(title="Population within 0.5 mile of Each Transit Stn", 
       subtitle="Boston (Suffolk County), MA", 
       caption="Data: U.S. Census Bureau, MBTA") +
  guides(color = guide_legend("as.numeric(partVal)"), size = guide_legend("as.numeric(partVal)")) +
  theme_void()+
  theme(legend.position = "right")
```

### 5.2. Map of Median Rent

annotate each visualization appropriately: The station of highest median rent is  Massachusetts Avenue Station as well, whose value is $45,000.

```{r PopRentPlot_12-22_2, warning = FALSE, message = FALSE}
#PLOT for Median Rent 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=medGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal), size = as.numeric(partVal), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="D")+
  scale_size_continuous(range=c(2,18))+
  labs(title="Median Rent within 0.5 mile of Each Transit Stn", 
       subtitle="Boston (Suffolk County), MA", 
       caption="Data: U.S. Census Bureau, MBTA") +
  guides(color = guide_legend("as.numeric(partVal)"), size = guide_legend("as.numeric(partVal)")) +
  theme_void()+
  theme(legend.position = "right")
```

## 6.

```{r medRent_Dist_Relation, warning = FALSE, message = FALSE}
allTracts.rings <-
  st_join(
    st_centroid(dplyr::select(allTracts, GEOID, year)),
    multipleRingBuffer(st_union(MBTA_NODE), 14484, 805)) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year),
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance=distance/1609)

ggplot() +
  geom_sf(data=allTracts.rings, aes(fill=distance))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", color='white', size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Distance to Subway Stns, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe white dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```

```{r medRent_Dist_Relation2, warning = FALSE, message = FALSE}
MedRent_sum22 <- allTracts.rings %>%
  filter(year %in% "2022", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2022)

MedRent_sum12 <- allTracts.rings %>%
  filter(year %in% "2012", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2012)

MedRent_sum <- rbind(MedRent_sum22,MedRent_sum12)


#MedRent_mean <- aggregate(MedRent ~ distance, data = allTracts.rings, FUN = mean, na.rm=TRUE, na.action=NULL)
MedRent_sum$year <- as.factor(MedRent_sum$year)
ggplot(MedRent_sum, aes(x = distance, y = mean_value, group=year, color=year)) +
  geom_line() +
  scale_x_continuous(limits = c(0.5, 3.6), breaks = seq(0.5, 3.6, by = 0.5)) +
  scale_y_continuous(limits = c(0, 2000), breaks = seq(0, 2000, by = 200)) +
  scale_color_manual(values = c("2022" = "red", "2012" = "blue"))+
labs(
    title = "Median Monthly Rent as a Function of Distance from Subway Stns",
    subtitle = "Boston (Suffolk County), MA",
    caption = "Data: U.S. Census Bureau, MBTA",
    x="Distance from Subway Stns (Miles)", 
    y="Median Monthly Rent ($)",
    color = "Year") +
  theme_minimal()
```

# Conclusion:

provide brief policy-relevant conclusions about willingness to pay for
TOD and how that might vary in the study area TOD에 대한 지불 의사와
연구 지역에서 이것이 어떻게 다를 수 있는지에 대한 간단한 정책 관련
결론을 제공합니다. 전체적으로는 지하철역에서부터 거리가 가까운
지역일수록 월세가 높아지는 것으로 나타났다. 특히 1.5마일까지는
지하철역에서부터 거리가 가까울수록 월세가 오르는 경향이 더 컸다. 다만
지하철에서부터 거리가 2마일 이상인 지역들 간에는 월세 차이가 없었다.
공간분석 결과로부터 지하철에서부터 거리가 1.5마일 이하인 지역에 대하여
TOD 정책이 임대료를 높인 것인지, 임대료가 높아서 TOD 정책을 실시하여야
하는지 인과관계를 직접적으로 알 수는 없었다. 다만 더 많은 사람들이
대중교통 접근을 높이고 지하철 이용률을 높이기 위해서는

TOD로 인하여 지하철역 주변 지역의 월세와 인구가 올랐는지, 아니면 인과
관계를 확인할 수는 없었지만,

,. The Massachusetts Bay Transportation Authority (MBTA) annotate each
visualization appropriately:

Your assignment is to prepare a policy brief for the local City Council
representatives in Boston (Suffolk County).

How certain can you be about your conclusions given some of the spatial
biases we’ve discussed?
