---
title: "HW2 - Boston TOD Analysis"
author: "Jun, Youngsang / yjmark@design.upenn.edu"
date: 'September 19, 2024'
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Brief Motivation

annotate each visualization appropriately:

# 1. Data Wrangling Work

Load Libraries

```{r}
# Install and Load Libraries
install.packages("mapview")
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(mapview)
library(dplyr)
library(viridis)

# Disable Scientific Notation
options(scipen=999)
options(tigris_class = "sf")

# Multiple
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

# Set Color Table
palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")

# Set ACS API Key
census_api_key("b2835f54d89a4499ba29829c908967b86765b345", overwrite = TRUE)
```

Wrangle American Community Survey (ACS) Data of Suffolk County (which
Boston seats of) 2012 and 2022

```{r}
# Get ACS Data of Suffolk County 2012 and 2022
tracts12 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2012, state="MA",
          county="Suffolk", geometry=TRUE) %>% 
  	
  st_transform('EPSG:26986') # Use NAD 83 CRS
tracts12 <- subset(tracts12, GEOID!='25025990101') # Error
tracts22 <-  
  get_acs(geography = "tract",
          variables = c("B01003_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), # Since B25026_001 is Total Population in Occupied Housing Units, used B01003_001E instead 
          year=2022, state="MA",
          county="Suffolk", geometry=TRUE) %>% 
  	
  st_transform('EPSG:26986') # Use NAD 83 CRS

# Transforming Long Data to Wide Data Using Spread
tracts12 <- 
  tracts12 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

tracts22 <- 
  tracts22 %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B01003_001, 
         Whites = B02001_002,
         FemaleBachelors = B15001_050, 
         MaleBachelors = B15001_009,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)

# Create Column that Calculates Percentage of Whites, Bachelors and Poverty
tracts12 <- 
  tracts12 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2012") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

tracts22 <- 
  tracts22 %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2022") %>%
  dplyr::select(-Whites,-FemaleBachelors,-MaleBachelors,-TotalPoverty)

# Row Bind of 2012 and 2022
allTracts <- rbind(tracts12,tracts22)
```

Wrangle Massachusetts Bay Transportation Authority (MBTA) Data

```{r}
# Get MBTA Data of Transit Line and Node
MBTA_ARC <- st_read('./MBTA_RT/MBTA_ARC.shp')
MBTA_NODE <- st_read('./MBTA_RT/MBTA_NODE.shp')

# Create a Column in MBTA_NODE to Distinguish Transfer Stations
MBTA_NODE <- MBTA_NODE %>%
  mutate(transferstation = ifelse(LINE == "BLUE", "BLUE",ifelse(LINE == "ORANGE", "ORANGE",ifelse(LINE == "RED", "RED",ifelse(LINE == "SILVER", "SILVER", ifelse(LINE == "GREEN", "GREEN","TRANSFER"))))))
```

Buffer Relating SEPTA Stops and Tracts and Set TOD/Non-TOD area

```{r buffer, include=FALSE, cache = TRUE}
stopBuffer <- st_buffer(MBTA_NODE, 805) # 805 meters which is equivalent to 0.5 miles
stopUnion <- st_union(st_buffer(MBTA_NODE, 805))
stopUnion <- st_make_valid(stopUnion) 
MBTABuffers <- 
  rbind(
     stopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     stopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))

ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, 
          aes(color = LINE), 
          show.legend = "line", size= 2) +
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E")) +
  geom_sf(data=MBTA_NODE,
          aes(color = transferstation), 
          show.legend = "point", size= 1)+
  scale_color_manual(values = c("#0C4C91", "#008445", "#F18C20", "#E12726", "#7C878E", "black"), labels = c("BLUE", "GREEN", "ORANGE", "RED", "SILVER", "TRANSFER STN")) +
  labs(title="MBTA", 
       subtitle="Boston, MA", 
       caption="Figure 2.5") +
  mapTheme()

# mapview(MBTA_ARC)+ mapview(MBTA_NODE)+ mapview(MBTABuffers)+ mapview(MBTABuffers)
```


```{r}
clip <- 
  st_intersection(buffer, tracts22) %>%
  dplyr::select(TotalPop) %>%
  mutate(Selection_Type = "Clip")
```

```{r}
# Do a spatial selection to see which tracts touch the buffer

# approach #1: sub-setting a spatial object with a spatial object using the '[' brackets.
selection1 <- 
  tracts22[buffer,] %>%
  dplyr::select(TotalPop) %>%
  mutate(Selection_Type = "Spatial Selection")

# approach #2: using `st_intersects` as a verbose way to do approach #1
selection2 <- tracts22[st_intersects(allTracts, buffer) %>% lengths > 0, ] %>%
  dplyr::select(TotalPop) %>%
  mutate(Selection_Type = "Spatial Selection")

# approach #3: use `st_join` to do a spatial join and remove the non-intersecting polygons
selection3 <- tracts22 %>% 
  st_join(buffer, join = st_intersects) %>% 
  filter(!is.na(Legend)) %>% 
  dplyr::select(TotalPop) %>%
  mutate(Selection_Type = "Spatial Intersects")
```

```{r}
selectCentroids22 <-
  st_centroid(tracts22)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts22, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2022")
selectCentroids22_union <- st_union(selectCentroids22)%>%
  st_sf() %>%
  mutate(year = "2022")
selectCentroids12 <-
  st_centroid(tracts12)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts12, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop) %>%
  mutate(year = "2012")
selectCentroids12_union <- st_union(selectCentroids12)%>%
  st_sf() %>%
  mutate(year = "2012")
intersections1222_union <- rbind(selectCentroids22_union, selectCentroids12_union)

NODE22 <- st_intersection(MBTA_NODE, tracts22) %>%
  mutate(year = "2022")
NODE12 <- st_intersection(MBTA_NODE, tracts12) %>%
  mutate(year = "2012")
MBTA_NODE_CLIP <-
  rbind(NODE22,NODE12)
```

```{r}
#intersections <- rbind(clip, selection1, selectCentroids22)

#ggplot() +
  #geom_sf(data=intersections, aes(fill = TotalPop)) +
  #geom_sf(data=MBTA_NODE, show.legend = "point") +
  #scale_fill_viridis_c() +
  #facet_wrap(~Selection_Type) + 
  #mapTheme()
```

# 2. Four small-multiple visualizations comparing four selected Census variables across time and space (TOD vs. non-TOD)

annotate each visualization appropriately:

```{r}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2012", MedRent * 1.2747, MedRent)) # Cumulative price change Inflation from 2012 to 2022 is	27.47%
```

```{r}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=MedRent))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Median Rent, 2012-2022",
    subtitle = "Real dollars; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```
```{r}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctWhite))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Percentage of White, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```
```{r}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctBachelors))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Percentage of Bachelors, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```
```{r}
ggplot() +
  geom_sf(data=allTracts.group, aes(fill=pctPoverty))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Percentage of Poverty, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe black dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()
```




```{r}
allTracts.Summary <- 
  st_drop_geometry(allTracts.group) %>%
  group_by(year, TOD) %>%
  summarize(Rent = mean(MedRent, na.rm = T),
            Population = mean(TotalPop, na.rm = T),
            Percent_White = mean(pctWhite, na.rm = T),
            Percent_Bach = mean(pctBachelors, na.rm = T),
            Percent_Poverty = mean(pctPoverty, na.rm = T))
```

# 3. Grouped Bar Plot

annotate each visualization appropriately:

```{r}
allTracts.Summary %>%
  gather(Variable, Value, -year, -TOD) %>%
  ggplot(aes(year, Value, fill= TOD ))+
    geom_bar(stat="identity", position="dodge")+
    facet_wrap(~Variable, scales="free", ncol=5)+
    scale_fill_manual(values=c("#bae4bc","#0868ac"))+
    labs(title="Indicator Differences across TIme and Space")+
    plotTheme() + theme(legend.position="bottom")
```

# 4. Table

annotate each visualization appropriately:

```{r}
allTracts.Summary %>%
  unite(year.TOD, year, TOD, sep = ": ", remove=T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Value=round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kable() %>%
    kable_styling()
```

# 5.a) A Graduated Symbol Map of Population within 0.5 mile of Each Transit Stn

annotate each visualization appropriately::

```{r}
# 트랙마다 인구밀도(=인구/면적) 계산
allTracts_Dens <- allTracts %>%
  mutate(tractArea=st_area(geometry)) %>%
  mutate(popDens=TotalPop/tractArea) %>%
  mutate(medDens=MedRent/tractArea)

# 역마다 버퍼 동그라미 면적 계산
stopBuffer <- stopBuffer %>%
  mutate(bufferArea=st_area(geometry)) 

# stopBuffer12 <- stopBuffer 

# Intersect (To Cut tracts22_Dens by stopBuffer12, 
# Loop through each stop and intersect with tracts22_Dens/tracts12_Dens)
stop_results <- data.frame()
hap <- data.frame()
for(i in 1:nrow(stopBuffer)) {
  stop_intersect <- st_intersection(stopBuffer[i,], allTracts_Dens) %>% 
    mutate(partArea=st_area(geometry)) %>% 
    mutate(partPop = partArea*popDens) %>% 
    mutate(partMed = partArea*medDens)
  stop_results <- rbind(stop_results, stop_intersect)
  print(i)
}

hapPop22 <- stop_results %>%
  filter(year %in% "2022") 
hapMed22 <- stop_results %>%
  filter(year %in% "2022") 
hapPop12 <- stop_results %>%
  filter(year %in% "2012")
hapMed12 <- stop_results %>%
  filter(year %in% "2012") 

a <- aggregate(hapPop22$partPop, by=list(hapPop22$STATION), FUN=sum, na.rm=TRUE, na.action=NULL) 
colnames(a) <- c("STATION","partVal") 
a <- a %>%
  mutate(year="2022") %>%
  mutate(classification="Population by Subway Stn")
  
b <-aggregate(hapMed22$partMed, by=list(hapMed22$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(b) <- c("STATION","partVal")
b <- b %>%
  mutate(year="2022") %>%
  mutate(classification="Median Rent by Subway Stn")

c <-aggregate(hapPop12$partPop, by=list(hapPop12$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(c) <- c("STATION","partVal") 
c <- c %>%
  mutate(year="2012") %>%
  mutate(classification="Population by Subway Stn")

d <-  aggregate(hapMed12$partMed, by=list(hapMed12$STATION), FUN=sum, na.rm=TRUE, na.action=NULL)
colnames(d) <- c("STATION","partVal")
d <- d %>%
  mutate(year="2012") %>%
  mutate(classification="Median Rent by Subway Stn")

popGraph <- rbind(a,c)
popGraph <- left_join(popGraph,MBTA_NODE, by="STATION")
medGraph <- rbind(b,d) 
medGraph <- left_join(medGraph,MBTA_NODE, by="STATION")
#PLOT for Population 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=popGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal), size = as.numeric(partVal), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="C")+
  scale_size_continuous(range=c(2,18))+
  labs(title="Population within 0.5 mile of Each Transit Stn", 
       subtitle="Boston, MA", 
       caption="Figure 2.5") +
  guides(color = guide_legend("as.numeric(partVal)"), size = guide_legend("as.numeric(partVal)")) +
  theme_void()+
  theme(legend.position = "right")


```

# 5.b) A Graduated Symbol Map of Rent within 0.5 mile of Each Transit Stn

annotate each visualization appropriately:

```{r}
#PLOT for Median Rent 2012-2022
ggplot() + 
  geom_sf(data=st_union(tracts22)) +
  geom_sf(data=MBTABuffers, alpha=0.5, fill="#E5E7E9", color = "gray") +
  geom_sf(data=MBTA_ARC, size= 2) +
  geom_point(
    data=medGraph,
    alpha=0.5,
    aes(color=as.numeric(partVal), size = as.numeric(partVal), geometry = geometry),
    stat = "sf_coordinates")+
  facet_wrap(~year)+
  scale_color_viridis(option="D")+
  scale_size_continuous(range=c(2,18))+
  labs(title="Median Rent within 0.5 mile of Each Transit Stn", 
       subtitle="Boston, MA", 
       caption="Figure 2.5") +
  guides(color = guide_legend("as.numeric(partVal)"), size = guide_legend("as.numeric(partVal)")) +
  theme_void()+
  theme(legend.position = "right")
```

# 6.
```{r}
allTracts.rings <-
  st_join(
    st_centroid(dplyr::select(allTracts, GEOID, year)),
    multipleRingBuffer(st_union(MBTA_NODE), 14484, 805)) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year),
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance=distance/1609)

ggplot() +
  geom_sf(data=allTracts.rings, aes(fill=distance))+
  geom_sf(data=intersections1222_union, fill='transparent', color='red', lwd = 1)+
    geom_sf(data=MBTA_NODE_CLIP, 
          show.legend = "point", color='white', size= 1)+
  facet_wrap(~year)+
  scale_fill_viridis()+
  labs(
    title = "Distance to Subway Stns, 2012-2022",
    subtitle = "0.0-1.0; The red border denotes areas close to subway stns;\nThe white dots denote subway stns",
    caption = "Data: U.S. Census Bureau, MBTA") +
  theme_void()

MedRent_sum22 <- allTracts.rings %>%
  filter(year %in% "2022", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2022)

MedRent_sum12 <- allTracts.rings %>%
  filter(year %in% "2012", !is.na(MedRent)) %>%
  group_by(distance) %>%
  summarize(mean_value = mean(MedRent)) %>%
  mutate(year=2012)

MedRent_sum <- rbind(MedRent_sum22,MedRent_sum12)


#MedRent_mean <- aggregate(MedRent ~ distance, data = allTracts.rings, FUN = mean, na.rm=TRUE, na.action=NULL)
MedRent_sum$year <- as.factor(MedRent_sum$year)
ggplot(MedRent_sum, aes(x = distance, y = mean_value, group=year, color=year)) +
  geom_line() +
  scale_x_continuous(limits = c(0.5, 3.6), breaks = seq(0.5, 3.6, by = 0.5)) +
  scale_y_continuous(limits = c(0, 2000), breaks = seq(0, 2000, by = 200)) +
  scale_color_manual(values = c("2022" = "red", "2012" = "blue"))+
labs(
    title = "Rent as a function of distance to Subway Stns",
    subtitle = "Boston, MA",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Miles", 
    y="Median Monthly Rent ($)",
    color = "Year") +
  theme_minimal()
```

# Conclusion:

provide brief policy-relevant conclusions about willingness to pay for
TOD and how that might vary in the study area
